---
output: html_document
editor_options:
  chunk_output_type: console
---
# Loading Data
```{r}
#install.packages("imputeTS")
library(tidyverse)
library(fpp3)
library(magrittr)
library(imputeTS)
rm(list = ls())
data <- read_delim("BaselMessungen.csv", delim = ";")
data %<>% as_tsibble() %<>%
  select(-c("Wasserstand","Pegel"))
```

data %>% arrange(desc(Zeitstempel)) %>% mutate(diff = lag(Zeitstempel) - Zeitstempel)%>% group_by(diff) %>% summarise(counter = n())
plot(data$Zeitstempel, data$Abflussmenge)
plot(data$Zeitstempel, data$Pegel)
plot(data$Zeitstempel, data$Wasserstand)
```

# Raw AR, MA, ARMA and ARIMA Functions

```{r}
my_MA <- function(ts, q, thetas){
  # verify equal length of q and thetas
  if(q+1!=length(thetas)){errorCondition("Inputs to my_MA are inapproriate.")}
  # thetas need to contain 1 as first value 
  # as the current noise innovation is not scaled => multiplied with 1
  out <- rep(NA, length(ts))
  
  for(i in seq_along(ts)){
    if(i-q<1){
      out[i] <- NA
      
    }else{
      out[i] <- sum(thetas * ts[i:i-q])
  }
  }
  return(out)  
}
my_AR <- function(ts, q, phis){
  # verify equal length of q and thetas
  if(q!=length(thetas)){errorCondition("Inputs to my_AR are inapproriate.")}
  # thetas need to contain 1 as first value 
  # as the current noise innovation is not scaled => multiplied with 1
  out <- rep(NA, length(ts))
  
  for(i in seq_along(ts)){
    if(i-q<1){
      out[i] <- ts[i]
      
    }else{
      out[i] <- sum(phis * out[seq(i-q,i-1,1)]) + ts[i]
    }
    out
  }
  return(out)  
}


# manual testing code
# ts <- c(1:10)
# q <- 3
# thetas <- c(1, 0.8, 0.9, 0.4)
# phis <- c(0.8, 0.9, 0.4)
# usual testing
# my_MA(
#   c(1:10),
#   3,
#   c(1, 0.8, 0.9, 0.4))
set.seed(123)
aux_ts <- rnorm(100)
ar <- my_AR(
  aux_ts,
  1,
  c(0.6))
aux_ts <- tsibble(ts = ts, index = c(1:length(ts)))
aux_ts %>% model(
  AR1 = ARIMA(ts ~ pdq(1, 0, 0), )
)
plot(c(1:length(ar)), ar, type ="l")

# First Analysis
```{r}
length(data$Zeitstempel)
length(data$Abflussmenge)
class(difference(data$Zeitstempel,1))
class(data$Zeitstempel)
class(data$Abflussmenge)

#data overview
plot(data$Zeitstempel,data$Abflussmenge, type="l")
ggplot(data=data)+
  geom_line(aes(x=Zeitstempel,y=Abflussmenge))
interval_pull(data$Zeitstempel)
summary(data)

#missing values----

#see missing values
sum(is.na(data$Abflussmenge))
sum(is.na(data$Abflussmenge))/length(data$Zeitstempel)

which(as.difftime(5, units = "mins") != difference(data$Zeitstempel,1))
na_index <- which(is.na(data$Abflussmenge)== TRUE)
View(data[na_index,])

#identify / fill gaps
has_gaps(data)
count_gaps(data)
scan_gaps(data)

#total missing values visualized
data %>% drop_na() %>% count_gaps() %>% mutate(val="Abflussmenge") %>%
  ggplot(aes(x = val, colour = val)) +
  geom_linerange(aes(ymin = .from, ymax = .to)) +
  geom_point(aes(y = .from)) +
  geom_point(aes(y = .to)) +
  coord_flip() +
  theme(legend.position = "bottom")

data %>% drop_na() %>% count_gaps() %>% count(.n)
data %>% drop_na() %>% count_gaps() %>% select(".n") %>%
  sum()

```
#Filling Gaps
```{r}

data %<>% fill_gaps( .full = TRUE)
ggplot_na_distribution(data)
ggplot_na_intervals(data)
ggplot_na_gapsize(data)
data_kalman <- na_kalman(data)
has_gaps(data_kalman)
ggplot_na_imputations(data, data_kalman)

```
#decomposition of time series with stl
```{r}
#aus fpp3
decomp<- data_kalman %>%
  model(stl = STL(Abflussmenge)) %>%
  components()

decomp %>%
  as_tsibble() %>%
  autoplot(Abflussmenge, colour="gray") +
  geom_line(mapping= aes(y=trend), colour = "#D55E00") +
  labs(
    y = "flow rate",
    x = "time",
    title = "Flow rate of the Rhine in Basel")

autoplot(decomp)

#is heteroscedastic
remainder <- decomp$remainder


acf(remainder)
```

# Raw Auto-Corr
```{r}
t = 0:300
y_stationary <- rnorm(length(t),mean=1,sd=1) # the stationary time series (ts)
plot(t, y_stationary, "l")

my_autocorr <- function(ts, q){
    aux <- tibble(counter = as.integer(), autocorr = as.double())
    for (i in seq_along(c(1:q))){
    table <-
        tibble(
            ts = ts,
            lagged_ts = lag(ts, i)) %>%
        drop_na() %>%
        summarise(
            autocorr = cov(ts, lagged_ts)/(sd(ts) * sd(lagged_ts))
            )
    out <- table %>% pull(autocorr)
    aux <- aux %>% complete(counter = i, autocorr = out)
    }
    return(aux)
}

test <- acf(y_stationary, 7)
test$acf[-1]
my_autocorr(y_stationary, 7) %>% arrange(counter)  %>% pull(autocorr)
((test$acf[-1] - my_autocorr(y_stationary, 7) %>% arrange(counter)  %>% pull(autocorr))/test$acf[-1]) %>% mean()

```
works fairly accurate — relative error 1%

# Package Functions
```{r}
library(fpp3)
```
